\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-05-17 redaction/dhp_sept2016.tex
comments_Table1

Table \ref{tab:summary_results} calls for two types of comments. Starting with pure statistical concerns, it appears that mean values are systematically higher than medians - by around one percentage point. This is no surprise at all: once again, our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median. Besides, one can also note that estimated total costs are also systematically higher than observed costs by 0.5 to 1 percentage point. This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - section 4 will go into more details of these compositions effects.

Coming now to interpretation, Table \ref{tab:summary_results} shows the magnitude of overall transport costs is sizeable. Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping. Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable. Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%. These values are respectively equal to 1.8\% and 2.5\% for air transport. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward. In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}

>>>>

Table \ref{tab:summary_results} calls for two types of comments. Starting with pure statistical concerns, it appears that mean values are systematically higher than medians - by around one percentage point. This is no surprise at all: once again, our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median. Besides, one can also note that estimated total costs are also systematically higher than observed costs by 0.5 to 1 percentage point. This is due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - section 4 will go into more details of these compositions effects. Furthermore, Table \ref{tab:summary_results} reports values/percentages, whereas all our estimates are based on log-linearizations. Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values. Descriptive statistics showing that it is indeed the case available upon request to the authors.

Coming now to interpretation, Table \ref{tab:summary_results} shows the magnitude of overall transport costs is sizeable. Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping. Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable. Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%. These values are respectively equal to 1.8\% and 2.5\% for air transport. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward. In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.
